%%\documentstyle[twocolumn,psfig]{article}
\documentstyle[11pt,psfig]{book}
%%\topmargin=-.32in
%%\headsep=.4in
%%\textheight=8.5in
%%\textwidth=6in
%%\oddsidemargin=.15in
%%\evensidemargin=.25in
\pagestyle{empty}
\begin{document}

\def\roadrunner{{\em Roadrunner}}
\def\rrfs{{\tt rrfs}}


\input{bootstrap}
\input{memory}
\input{interrupts}


\chapter{Time}

The kernel tracks the progress of time.  The kernel uses a timer
interrupt to indicate the passage of time and keeps the time of
day for use by the kernel and application programs.  The kernel
also implements time-slicing of the CPU to implement
multi-processing.


\section{Time of Day}


\section{Time-Slicing}


\chapter{Processes}

Processes represent executing application programs.  Each process
requires a collection of hardware and kernel resources to execute
an application program.  The \roadrunner\  kernel implements
fully preemptible multi-processing, i.e. any process can be
interrupted at any time.


\section{Process States}


\section{Process Queues}


\chapter{Mutexes}

Mutexes are used within the kernel to synchronize simultaneous
accesses by multiple processes to certain hardware and kernel
elements.  Processes that want to access a shared resource may
be forced to wait in a process queue.


\chapter{Events}

Events are analogous to interrupts.  The difference is that
handling of events is done using processes rather than interrupt
service routines.  Processes are placed in a process list and
all are scheduled for execution when a specified event occurs.


\chapter{Device Drivers}

Input/Output (I/O) is handled by peripheral device hardware.
Device drivers represent the kernel elements that manage I/O to
and from hardware devices.  The kernel provides a uniform view of
all hardware devices, despite their differing I/O characteristics.


\chapter{File Systems}

There are various file system formats in which data can be
stored and presented.  The kernel has the ability to use multiple
file system format simultaneously within a single file system
name space.  A generic Application Programming Interface (API)
is provided to access data regardless of the underlying file
system format.


\section{The System File System ({\tt sysfs})}


\section{The Device File System ({\tt devfs})}


\input{rrfs}


\chapter{Lexical Analysis}

Lexical analysis provides capabilities for complex ASCII string
manipulation.  Capabilities are provided to decompose a string
into a sequence of arguments, to provide argument typing, and
to convert number strings into numeric values.


\chapter{Pathnames}

Files represent a generic mechanism for storing and presenting
data.  Files are stored in a hierarchical file system and
pathnames are used to locate files in a file system.  Routines
are provided that allow construction and decomposition of
pathnames.


\chapter{Files}

File represent the basic mechanism for storing and presenting
data.  The kernel implements a POSIX interface for access to
all file data.


\chapter{Program Loading}

Programs are loaded via the file system interface into memory
for execution.  \roadrunner\  uses the Executable and Linkable
Format (ELF) executable file format for application programs.


\chapter{System Calls}

System calls are used by application programs or libraries to
invoke kernel services.  The \roadrunner\  system call interface
is based on the POSIX standards.


\end{document}
